---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "Training in ade4 in R - Module I: Basic methods"
```

# LausanneBasic

[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/sdray/LausanneBasic/main?urlpath=rstudio)

This repository contains the material for the course *`r Title`*. 


## Contents

The directory contains the material for the first four sessions (course and practical). The last session is designed for open discussion and personal work. The structure of this compendium is as follows:

  - [:file\_folder: course](/course): Source (R Markdown) to reproduce the slides of the course. It contains also rendered versions (html and pdf documents) suitable for reading (the code is replaced by figures and tables in these files)
  - [:file\_folder: practical](/practical): Source and rendered versions of the slides. Corrections of practical are provided as R codes to reproduce the different analyses

## Program
* **Session 1**: Welcome, presentation of participants. Introduction, packages installation. Theory and 3D explanation of multivariate methods.
    - Course [[html](course/session1/session1.html)|[pdf](course/session1/session1.pdf)|[Rmd](course/session1/session1.Rmd)]
    - Practical [Rmd|html|pdf]
* **Session 2**: Analysis of environmental data: Principal Component Analysis.
    - Course [Rmd|html|pdf]
    - Practical [Rmd|html|pdf]
* **Session 3**: Analysis of environmental data: Multiple Correspondence Analysis. Hill-Smith Analysis
    - Course [Rmd|html|pdf]
    - Practical [Rmd|html|pdf]
* **Session 4**:Analysis of species data: Correspondence Analysis and Principal Component Analysis. Principal Coordinates Analysis.
    - Course [Rmd|html|pdf]
    - Practical [Rmd|html|pdf]
* **Session 5**: Examples of use of multivariate analysis methods with ade4 in R. Group work and discussions on data sets presented by participants.


## How to run in your broswer or download and run locally

This course has been developed using the statistical programming
language R. To work with the compendium, you will need
installed on your computer the [R software](https://cloud.r-project.org/)
itself and optionally [RStudio Desktop](https://rstudio.com/products/rstudio/download/).

You can download the compendium as a zip from from this URL:
[master.zip](/archive/master.zip). After unzipping:
- open the `.Rproj` file in RStudio
- source `make.R` file to compile the different documents of the course.

### How to cite

Please cite this document as:

> Dray, S. (`r format(Sys.Date(), "%Y")`). `r Title` . Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>


### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. 

### Session Information
```{r}
utils::sessionInfo()
```
