---
title: "Training in ade4 in R - Module I: Basic methods"
subtitle: "Practical 1: understanding multivariate methods in 3D"  
author: "Stéphane Dray"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css:  [custom.css, xaringan-themer.css] 
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
    includes:
            before_body: preamble.tex
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(tidy=TRUE, tidy.opts=list(width.cutoff=50), fig.height= 4, fig.width=4, dev='svg', fig.align="center", fig.path = "fig/")
```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)
style_duo(primary_color = "#1F4257", secondary_color = "#F97B64", table_row_even_background_color = "#1F4257")
```

```{r rgl-plot, include = FALSE}
library(rgl)
knitr::knit_hooks$set(webgl = hook_webgl)
```
---
# Introduction

In this practical we will try to understand the main principles of multivariate methods by interactive representation in 3 dimensions using the `rgl` package

First, load the required packages:

```{r load-pkg, message = FALSE}
library(ade4)
library(rgl)
```

and then the data set `doubs` available in `ade4`

```{r}
data(doubs)
```
---
# Data

The data set contains information on 30 sites sampled on the Doubs river:

- Measurments of 11 environmental variables in `$env`
- Abundances of 27 fish species in `$fish`
- Spatial coordinates of the sites in `$xy`

More details? Try `?doubs`

Select 3 environmental variables for this practical and define a vector to color the different sites:

```{r}
tab <- doubs$env[,c(3,8,10)]
color <- rainbow(30, alpha = 0.5)
```
---
# Representing data: space of individuals

Load some utilities functions available in the R directory and plot the original data:


```{r plot-sim, eval=FALSE}
source("../../R/3D-utils.R")
plot3d(tab, type = "s", col = color)
```

.pull-left[
```{r plot-sim-out, message = FALSE, ref.label="plot-sim", echo=FALSE, fig.dim=c(4, 4), webgl=TRUE}
```
]

.pull-right[
* Our aim it to rotate the cloud to have the best viewpoint on the data.
* How can you charachterize this viewpoint?
* Look at scales!
]
---
This representation does not use the same scaling for variables due to the differences in ranges:

```{r, results='hide'}
summary(tab)
```

The same representation with iso-scaling:
```{r, eval = FALSE}
lims <- 1.1 * c(min(tab),max(tab))
plot3d(tab, type = "s", col = color, xlim = lims, ylim = lims, zlim = lims)
```

* What is the problem?

* How to solve it?
---
Data are in different units. Main variations are mainly driven by differences in variances. Standardization is required.

Center the data and see which geometric operation is induced:

```{r, eval = FALSE}
## centring
tab.c <- scale(tab, center = TRUE, scale = FALSE)
myplot3d(tab.c, colpoints = color, colaxes = 'black')
summary(tab.c)
```

Scale the data and see which geometric operation is induced:
```{r, eval = FALSE}
## scaling
tab.sc <- scale(tab, center = TRUE, scale = TRUE)
myplot3d(tab.sc, colpoints = color, colaxes = 'black')
```

Rotate and find the best viewpoint.
---
# Principal component analysis

Draw the principal axes

```{r, eval = FALSE}
pca <- dudi.pca(tab, nf = 3, scannf = FALSE) 
addvar3d(t(pca$c1), col = 'red')
```

and project the sites on the principal axes

```{r, eval = FALSE}
## first PCA axis
for(i in 1:nrow(pca$tab))
  segments3d(rbind(pca$c1[,1] * pca$li[i,1], tab.sc[i,]), col = color[i], lty = 3)

## second PCA axis
for(i in 1:nrow(pca$tab))
    segments3d(rbind(pca$c1[,2] * pca$li[i,2], tab.sc[i,]), col = color[i])
```
---
Lastly, we can use the principal axes as a new system of coordinates and represent the data in this new system

```{r, eval = FALSE}
open3d()
myplot3d(pca$li, colpoints = color, colaxes = 'black')
addvar3d((pca$c1), col = 'red')
```
---
# Space of variables

The same approach can be used to search for the best representation of variables. 

```{r, eval = FALSE}
tab2 <- doubs$env[c(1,11,27),]
tab2.sc <- t(scale(doubs$env[c(1,11,27),]))
color.var <- rainbow(11)
myplot3d2(tab2.sc, colarrows = color.var, colaxes = 'black')

## add PCA axes
pca2 <- dudi.pca(tab2, nf = 3, scannf = FALSE) 
addvar3d(t(pca2$l1), col = 'red')
```

.center[
[Go back to course 1](../../course/session1/session1.html#12)

]
---
# PCA by hand

Use matrix algebra and the functions `cor`, `%*%` and `eigen` to recompute the outputs of PCA by hand

```{r, echo = FALSE, eval = FALSE}
eig <- eigen(cor(tab))
all.equal(eig$values, pca$eig)
all.equal(eig$vectors[,1:3], as.matrix(pca$c1), check.attributes = FALSE)
row.coord <- as.matrix(pca$tab) %*% eig$vectors[,1:3]
all.equal(as.matrix(pca$li), row.coord, check.attributes = FALSE)
```

**Bonus**: Perform the singular value decomposition (function `svd`) of the standardized table (`scalewt(tab)`) and compare to the outputs of PCA


.center[
[Go back to course 1](../../course/session1/session1.html#18)

]
