---
title: "Training in ade4 in R - Module I: Basic methods"
subtitle: "Correspondence analysis"  
author: "Stéphane Dray"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css:  [custom.css, xaringan-themer.css] 
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
    includes:
            before_body: preamble.tex
params:
  correction: FALSE
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(tidy=TRUE, tidy.opts=list(width.cutoff=50), fig.height= 4, fig.width=4, dev='svg', fig.align="center", fig.path = "fig/")
```

```{r correction, include = FALSE}
res.correction <- ifelse(params$correction, "markup", "hide")
```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)
style_duo(primary_color = "#1F4257", secondary_color = "#F97B64", table_row_even_background_color = "#1F4257")
```
---
# Data

We will analyze the `doubs` data set (see `?doubs`)

```{r}
library(ade4)
library(adegraphics)
data(doubs)
names(doubs)
names(doubs$fish)
```
---
# Correspondence Analysis

* Perform CA
* Display the barplot of eigenvalues

```{r, out.width="30%", include = params$correction, echo = params$correction, results=res.correction}
coa1 <- dudi.coa(doubs$fish, scannf = FALSE)
screeplot(coa1, main = " ", xlab = " ")
```
---
# Inertia statistics

* Compute the percentage of variation explained bu the first PCA axes

```{r, include = params$correction, echo = params$correction, results=res.correction}
summary(coa1)
```

---
# Graphical representation of CA results

* Plot the results using the `biplot` function 

```{r, fig.dim=c(3.5, 3.5), include = params$correction, echo = params$correction, results=res.correction}
biplot(coa1)
```
---
# CA scores on the geographical map

* Draw maps of CA scores on the first two axes
* Interpret the maps to describe how the fish communities vary along the river

```{r, fig.dim=c(4,2), include = params$correction, echo = params$correction, results=res.correction}
mypal <- colorRampPalette(c("#EDF8FB", "#006D2C"))
s.value(doubs$xy, coa1$li[, 1:2], pgrid.draw = FALSE, porigin.draw = FALSE,
    method = "size", symbol = "circle", col = mypal(2),
    ppoints.cex = 1)
```
---
# Principal Component Analysis

PCA can also be applied on the abundance table. Perform PCAon `doubs$fish` table. Should we scale or not?

```{r, fig.dim=c(3.5, 3.5), include = params$correction, echo = params$correction, results=res.correction}
pca1 <- dudi.pca(doubs$fish, scale = FALSE, scannf = FALSE)
biplot(pca1)
```
---
# PCA vs CA

Compare the biplots of CA and PCA

.pull-left[
```{r, include = params$correction, echo = params$correction, results=res.correction}
biplot(coa1)
```
]

.pull-right[
```{r, include = params$correction, echo = params$correction, results=res.correction}
biplot(pca1)
```
]
---
# Principal Coordinates Analysis

* Compute Jaccard distances between sites 
* Perform principal coordinates analysis and display the ordination of sites

```{r, include = params$correction, echo = params$correction, results=res.correction}
dis <- dist.binary(doubs$fish, method = 1)
pco1 <- dudi.pco(dis, scannf = FALSE)
s.label(pco1$li)
```
---
# PCA and PCoA

* Compare the results of PCA and those of PCoA applied on Euclidean distance (function  `dist`)

```{r, include = params$correction, echo = params$correction, results=res.correction}
dis2 <- dist(doubs$fish)
pco2 <- dudi.pco(dis2, scannf = FALSE)
```

.pull-left[
```{r, include = params$correction, echo = params$correction, results=res.correction}
head(pca1$eig, 2)
head(pca1$li)
```
]

.pull-right[
```{r, include = params$correction, echo = params$correction, results=res.correction}
head(pco2$eig, 2)
head(pco2$li)
```
]


